---
title: "Group S Final Report"
author: "Sarah Fatihi, Sophie Diop, James Street" 
output: pdf_document
date: "2024-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 6)

#Libraries 
library(tidyverse)
library(mosaic)
library(broom)
library(rvest)
library(GGally)
library(gridExtra)

#Data Sets 
players <- readRDS("data/player_data.rds")
```

## Abstract

## Background and Meaning 

## Methods

## Results 

### Univariate Descriptive Stats and Figures
```{r, echo=FALSE, results='hide'}
#UNIVARIATE ANALYSIS
#Quantitative Variables
ggplot(players, aes(x = value_eur)) +
  geom_density(fill = "skyblue", color = "blue", alpha = 0.5) +
  labs(title = "Distribution of Value (in Euros)", x = "Players' Values (in euros)")

ev1 <- ggplot(players, aes(x = pace)) +
  geom_density(fill = "skyblue", color = "blue", alpha = 0.5)+
  labs(title = "Distribution of Pace", x = "Players' Pace Rating")

ev2 <- ggplot(players, aes(x = shooting)) +
  geom_density(fill = "skyblue", color = "blue", alpha = 0.5)+
  labs(title = "Distribution of Shooting", x = "Players' Shooting Rating")
favstats(~ shooting, data = players)

ev3 <- ggplot(players, aes(x = passing)) +
  geom_density(fill = "skyblue", color = "blue", alpha = 0.5)+
  labs(title = "Distribution of Passing", x = "Players' Passing Rating")

ev4 <- ggplot(players, aes(x = dribbling)) +
  geom_density(fill = "skyblue", color = "blue", alpha = 0.5)+
  labs(title = "Distribution of Dribbling", x = "Players' Dribbling Rating")

ev5 <- ggplot(players, aes(x = defending)) +
  geom_density(fill = "skyblue", color = "blue", alpha = 0.5)+
  labs(title = "Distribution of Defending", x = "Players' Defending Rating")

ev6 <- ggplot(players, aes(x = physic)) +
  geom_density(fill = "skyblue", color = "blue", alpha = 0.5)+
  labs(title = "Distribution of Physicality", x = "Players' Physicality Rating")

#Qualitative Variables
ggplot(players, aes(x = club_name)) +
  geom_bar(fill = "skyblue", color = "blue") +
  labs(title = "Distribution of Clubs", x = "Clubs") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#FAVSTATS
favstats(~ pace, data = players)
favstats(~ shooting, data = players)
favstats(~ passing, data = players)
favstats(~ dribbling, data = players)
favstats(~ defending, data = players)
favstats(~ physic, data = players)
```

```{r, echo=FALSE, results='hide'}
#DISPLAYED GRAPHS
grid.arrange(ev1, ev2, ev3, ev4, ev5, ev6, ncol=3) 
```

Pace: The ratings of players' pace is left-skewed, so we use the median of 71 as a measure of center, and the IQR of 13.75 (first quartile = 64.25, third quartile = 78) as a measure of spread. 

Shooting: The ratings of players' shooting is left-skewed, so we use the median of 61 as a measure of center, and the IQR of 24 (first quartile = 47, third quartile = 71) as a measure of spread. 

Passing: The ratings of players' shooting has a slight left-skew, so we use the median of 67 as a measure of center, and the IQR of 14 (first quartile = 59, third quartile = 73) as a measure of spread.

Dribbling: The ratings of players' dribbling is left-skewed, so we use the median of 72 as a measure of center, and the IQR of 12 (first quartile = 65, third quartile = 77) as a measure of spread.

Defending: The ratings of players' defending is left-skewed with a potential second peak, so we use the median of 66 as a measure of center, and the IQR of 29 (first quartile = 46, third quartile = 75) as a measure of spread.

Physicality: The ratings of players' physicality is left-skewed, so we use the median of 71 as a measure of center, and the IQR of 13 (first quartile = 63, third quartile = 76) as a measure of spread.

### Bivaraite Analysis

## Interpreting the Model 

## Conclusion 

